```{r}
library(kableExtra)
library(chillR)
library(tidyverse)
library(ggpmisc)
library(patchwork)
```

# Chapter 18 Plotting future Scenarios

Produce similar plots for the weather station you selected for earlier exercises.

As you may have noticed the analysed plots produced with the inbuilt chillR functions are not very clear. To produce clearer plots we can use ggplot, which has already a build in function in chillR.
For the plots I will use the scenario data that I produced recently

```{r}
# import the chill scenario data and partly store it as lists with the load_temperature_scenarios(). The lists can be used as input now. 

chill_hist_scenario_list <- load_temperature_scenarios("data",
                                                       "Dehesa_hist_chill_305_59")
actual_chill <- read_tab("data/Dehesa_observed_chill_305_59.csv")

chill_future_scenario_list <- load_temperature_scenarios("future_climate","Dehesa_futurechill_305_59")

# using the make_climate_scenario() function to create climate scenarios for plotting. The function uses climate metric data that is stored in lists. I use the list I just generated.

chills <- make_climate_scenario(
  chill_hist_scenario_list,
  caption = "Historic",
  historic_data = actual_chill,
  time_series = TRUE)

# filtering the data for the scenarios
SSPs <- c("ssp126", "ssp245", "ssp370", "ssp585")
Times <- c(2050, 2085)

list_ssp <- 
  strsplit(names(chill_future_scenario_list), '\\.') %>%
  map(2) %>%
  unlist()

list_gcm <-
  strsplit(names(chill_future_scenario_list), '\\.') %>%
  map(3) %>%
  unlist()

list_time <-
  strsplit(names(chill_future_scenario_list), '\\.') %>%
  map(4) %>%
  unlist()


for(SSP in SSPs)
  for(Time in Times)
    {
    
    # find all scenarios for the ssp and time
    chill <- chill_future_scenario_list[list_ssp == SSP & list_time == Time]
    names(chill) <- list_gcm[list_ssp == SSP & list_time == Time]
    if(SSP == "ssp126") SSPcaption <- "SSP1"
    if(SSP == "ssp245") SSPcaption <- "SSP2"
    if(SSP == "ssp370") SSPcaption <- "SSP3"
    if(SSP == "ssp585") SSPcaption <- "SSP5"    
    if(Time == "2050") Time_caption <- "2050"
    if(Time == "2085") Time_caption <- "2085"
    chills <- chill %>% 
      make_climate_scenario(
        caption = c(SSPcaption,
                    Time_caption),
        add_to = chills)
}
```

For ggplot using the data.frame format is recommended. Therefore I loop through our chill projection list and extract the data. First I do it for the historic data.
```{r}


for(nam in names(chills[[1]]$data))
  {
   # Extract the data frame.
   ch <- chills[[1]]$data[[nam]]
   # Add columns for the new information we have to add and fill them.
   ch[,"GCM"] <- "none"
   ch[,"SSP"] <- "none"
   ch[,"Year"] <- as.numeric(nam)
   
   # Now check if this is the first time we've gone through this loop.
   # If this is the first time, the ch data.frame becomes the output
   # object (past_simulated).
   # If it is not the first time ('else'), we add the current data.frame
   # to the 'past_simulated' object
  if(nam == names(chills[[1]]$data)[1])
    past_simulated <- ch else
      past_simulated <- rbind(past_simulated,
                              ch)
  }

# Add another column called 'Scenario' and label all rows as 'Historical' 
past_simulated["Scenario"] <- "Historical"

kable(head(past_simulated))  %>%
  kable_styling("striped", position = "left",font_size = 8)
```
To add the historic observations to my data I create a pointer.
```{r}
past_observed <- chills[[1]][["historic_data"]]
```
In the next step I loop through the future data and add it to the list.
```{r}
# Extract future data
for(i in 2:length(chills))
  for(nam in names(chills[[i]]$data))
    {ch <- chills[[i]]$data[[nam]]
     ch[,"GCM"] <- nam
     ch[,"SSP"] <- chills[[i]]$caption[1]
     ch[,"Year"] <- chills[[i]]$caption[2]
     if(i == 2 & nam == names(chills[[i]]$data)[1])
       future_data <- ch else
         future_data <- rbind(future_data,ch)
  }
```
Now I can plot the data. Therefore I combine multiple plots using the plot_layout()-function, that is contained in the patchwork package. To prevent axis scales due to variation in the data I set the scales using the range() function.

```{r}
metric <- "GDH"
axis_label <- "Heat (in GDH)"

# get extreme values for the axis scale

rng <- range(past_observed[[metric]],
             past_simulated[[metric]],
             future_data[[metric]]) 
```
Now I first produce a plot for the historic data. Therefore I generate box plots of the simulated historic data and combine them with a scatter plot of the observed historic data.

```{r}
past_plot <- ggplot() +
  geom_boxplot(data = past_simulated,
               aes_string("as.numeric(Year)", metric, group = "Year"),
               fill = "skyblue",
               width = 0.9) +  # Increased boxplot width
  scale_y_continuous(limits = c(0, round(rng[2] + rng[2]/10))) +
  labs(x = "Year", y = axis_label) +
  facet_grid(~ Scenario) +
  theme_bw(base_size = 15) +
  geom_point(data = past_observed,
             aes_string("End_year", metric),
             col = "blue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



past_plot
```

Next I produce the future scenarios for the years 2050 and 2085 and combine them.

```{r}
future_plot_list <- list()

time_points <- c(2050, 2085)

for(y in time_points)
{
  future_plot_list[[which(y == time_points)]] <-
    ggplot(data = future_data[which(future_data$Year==y),]) +
    geom_boxplot(aes_string("GCM",
                            metric,
                            fill="GCM")) +
    facet_wrap(vars(SSP), nrow = 1) +
    scale_x_discrete(labels = NULL,
                     expand = expansion(add = 1)) +
    scale_y_continuous(limits = c(0, 
                                  round(round(1.1*rng[2])))) +
    geom_text_npc(aes(npcx = "center",
                      npcy = "top", 
                      label = Year),
                  size = 5) +
    theme_bw(base_size = 15) +
    theme(axis.ticks.y = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          legend.position = "bottom",
          legend.margin = margin(0, 
                                 0, 
                                 0, 
                                 0, 
                                 "cm"),
          legend.background = element_rect(),
          strip.background = element_blank(),
          strip.text = element_text(face = "bold"),
          legend.box.spacing = unit(0, "cm"),
          plot.subtitle = element_text(
            hjust = 0.5,
            vjust = -1,
            size = 15 * 1.05,
            face = "bold")) 
}

future_plot_list
```
In the end I combine all produced plots.
```{r}
both_plots <- past_plot + future_plot_list

plot <- both_plots +
           plot_layout(guides = "collect",
                       widths = c(1,rep(2,length(future_plot_list))))

plot <- plot & theme(legend.position = "bottom",
                     legend.text = element_text(size=8),
                     legend.title = element_text(size=10),
                     axis.title.x = element_blank())

plot
```

I also have to produce plots for Chill Portions and Frost duration. For faster plotting of these and in the future I wrap up the code produced previously into a function

```{r}
plot_scenarios_gg <- function(past_observed,
                              past_simulated,
                              future_data,
                              metric,
                              axis_label,
                              time_points)
{
  rng <- range(past_observed[[metric]],
               past_simulated[[metric]],
               future_data[[metric]])  
  past_plot <- ggplot() +
    geom_boxplot(data = past_simulated,
                 aes_string("as.numeric(Year)",
                            metric,
                            group="Year"),
                 fill="skyblue") +
    scale_y_continuous(limits = c(0, 
                                  round(round(1.1*rng[2])))) +
    labs(x = "Year", y = axis_label) +
    facet_grid(~ Scenario) +
    theme_bw(base_size = 15) +  
    theme(strip.background = element_blank(),
          strip.text = element_text(face = "bold"),
          axis.text.x = element_text(angle=45, 
                                     hjust=1)) +
    geom_point(data = past_observed,
               aes_string("End_year",
                          metric),
               col="blue")
  
  future_plot_list <- list()
  
  for(y in time_points)
  {
    future_plot_list[[which(y == time_points)]] <-
      ggplot(data = future_data[which(future_data$Year==y),]) +
      geom_boxplot(aes_string("GCM", 
                              metric, 
                              fill="GCM")) +
      facet_wrap(vars(SSP), nrow = 1) +
      scale_x_discrete(labels = NULL,
                       expand = expansion(add = 1)) +
      scale_y_continuous(limits = c(0, 
                                    round(round(1.1*rng[2])))) +
      geom_text_npc(aes(npcx = "center",
                        npcy = "top",
                        label = Year),
                    size = 5) +
      theme_bw(base_size = 15) +
      theme(axis.ticks.y = element_blank(),
            axis.text = element_blank(),
            axis.title = element_blank(),
            legend.position = "bottom",
            legend.margin = margin(0,
                                   0, 
                                   0, 
                                   0, 
                                   "cm"),
            legend.background = element_rect(),
            strip.background = element_blank(),
            strip.text = element_text(face = "bold"),
            legend.box.spacing = unit(0, "cm"),
            plot.subtitle = element_text(hjust = 0.5,
                                         vjust = -1,
                                         size = 15 * 1.05,
                                         face = "bold")) 
  }
  
  plot <- (past_plot +
             future_plot_list +
             plot_layout(guides = "collect",
                         widths = c(1,rep(2,length(future_plot_list))))
           ) & theme(legend.position = "bottom",
                     legend.text = element_text(size = 8),
                     legend.title = element_text(size = 10),
                     axis.title.x=element_blank())
  plot
  
}
```

```{r}
plot_scenarios_gg(past_observed = past_observed,
                  past_simulated = past_simulated,
                  future_data = future_data,
                  metric = "Chill_Portions",
                  axis_label = "Chill (in Chill Portions)",
                  time_points = c(2050, 2085))
```

```{r}
plot_scenarios_gg(past_observed = past_observed,
                  past_simulated = past_simulated,
                  future_data = future_data,
                  metric = "Frost_H",
                  axis_label = "Frost duration (in hours)",
                  time_points = c(2050, 2085))
```

