
To delineate temperature response phases with PLS regression we can use the PLS_pheno()- tool in chillR. 
```{r}
library(chillR)
#as the first step we convert the first bloom dates to Julian dates

Roter_Boskop <- read_tab("data/Roter_Boskoop_bloom_1958_2019.csv") %>%
  select(Pheno_year, First_bloom) %>%
  mutate(Year = as.numeric(substr(First_bloom, 1, 4)),
         Month = as.numeric(substr(First_bloom, 5, 6)),
         Day = as.numeric(substr(First_bloom, 7, 8))) %>%
  make_JDay() %>%
  select(Pheno_year, JDay) %>%
  rename(Year = Pheno_year,
         pheno = JDay)
```
We also import the weather data and also assign Julian dates.
```{r}
Dehesa_temps <- read_tab("data/TMaxTMin1958-2019_patched.csv") %>%
  make_JDay
```
In the next step we conduct the PLS regression using the PLS_pheno function.
```{r}
PLS_results <- PLS_pheno(Dehesa_temps,
                         Roter_Boskop)


```
The function produces a list containing two data frames. The results are contained in the PLS_summary data.frame, which we can now plot using the plot_PLS()-function. 

```{r}
plot_PLS(PLS_results, "data/pls_roter_boskoop")
```




