```{r}
library(chillR)
library(devtools)
library(dormancyR)
library(matlab)
library(tidyverse)
library(ggplot2)
```
```{r}
library(colorRamps)
library(ggplot2)
library(patchwork)
```
# chill model comparison at the Dehesa San Francisco

There are various chill models that can be used for phenology analysis. To choose the right one it might be useful to compare those models. A useful statistic for that is Safe Winter Chill. Safe Winter Chill is the 10% quantile of chill distribution, corresponding to chill amount that is exceeded in 90% of all years. We can compute the Safe Winter Chill for various chill model and produce heat maps for comparison. As an excample we will do this for the Dehesa San Francisco

```{r}
# first I create lists of the models I use. The list hourly_models contains models that are based on hourly temperature data, the daily_models list contains models based on daily data
hourly_models <- list(Chilling_units = chilling_units,
     Low_chill = low_chill_model,
     Modified_Utah = modified_utah_model,
     North_Carolina = north_carolina_model,
     Positive_Utah = positive_utah_model,
     Chilling_Hours = Chilling_Hours,
     Utah_Chill_Units = Utah_Model,
     Chill_Portions = Dynamic_Model)

daily_models <- list(Rate_of_Chill = rate_of_chill,
                     Chill_Days = chill_days,
                     Exponential_Chill = exponential_chill,
                     # Triangular_Chill_Haninnen = triangular_chill_1,
                     Triangular_Chill_Legave = triangular_chill_2)

# next I store the models together in a data.frame
metrics <- c(names(daily_models),
             names(hourly_models))

model_labels = c("Rate of Chill",
                 "Chill Days",
                 "Exponential Chill",
                 # "Triangular Chill (Häninnen)",
                 "Triangular Chill (Legave)",
                 "Chilling Units",
                 "Low-Chill Chill Units",
                 "Modified Utah Chill Units",
                 "North Carolina Chill Units",
                 "Positive Utah Chill Units",
                 "Chilling Hours",
                 "Utah Chill Units",
                 "Chill Portions")
```

```{r}
data.frame(Metric = model_labels, 'Function name' = metrics)
```
For comparison I apply the models to the observed temperature data as well as to the generated historical and future scenarios
```{r}
#therefore I first load our temperature data

Dehesa_temps <- read_tab("Dehesa_temps.csv")

Temps <- load_temperature_scenarios("data",
                                    "Dehesa_hist_scenarios")

future_temps <- load_temperature_scenarios("future_climate","Dehesa_futuretemps")
```

Now I apply the models to the past temperature scenarios. For the models using hourly data we again use the tempResponse_daily_list() function. For the models using daily data we use the tempResponse_list_daily() function of the dormancyR package instead, which is the equivalent tool for daily data.
```{r}
Start_JDay <- 305
End_JDay <- 59

daily_models_past_scenarios <- 
  tempResponse_list_daily(Temps,
                          Start_JDay = Start_JDay,
                          End_JDay = End_JDay,
                          models=daily_models)

# when using the daily-function you have to deleve incomplete data manually

daily_models_past_scenarios <- lapply(
  daily_models_past_scenarios,
  function(x) x[which(x$Perc_complete>90),])

hourly_models_past_scenarios<-
  tempResponse_daily_list(Temps,
                          latitude = 37.8667,
                          Start_JDay = Start_JDay,
                          End_JDay = End_JDay,
                          models = hourly_models,
                          misstolerance = 10)

past_scenarios <- daily_models_past_scenarios

# after conducting the models to both list I merge them to a single data frame, using the lapply()tool.
past_scenarios <- lapply(
  names(past_scenarios),
  function(x)
    cbind(past_scenarios[[x]],
          hourly_models_past_scenarios[[x]][,names(hourly_models)]))

names(past_scenarios) <- names(daily_models_past_scenarios)

# the same workflow I repeat for the observed temperature data

daily_models_observed <- 
  tempResponse_daily(Dehesa_temps,
                     Start_JDay = Start_JDay,
                     End_JDay = End_JDay,
                     models = daily_models)

daily_models_observed <-
  daily_models_observed[which(daily_models_observed$Perc_complete>90),]

hourly_models_observed <- 
  tempResponse_daily_list(Dehesa_temps,
                          latitude=37.8667,
                          Start_JDay = Start_JDay,
                          End_JDay = End_JDay,
                          models = hourly_models,
                          misstolerance = 10)

past_observed <- cbind(
  daily_models_observed,
  hourly_models_observed[[1]][,names(hourly_models)])

# for later use I save the scenarios.

save_temperature_scenarios(past_scenarios,
                           "data/future_climate",
                           "Dehesa_multichill_305_59_historic")
write.csv(past_observed,
          "data/future_climate/Dehesa_multichill_305_59_observed.csv",
          row.names=FALSE)
```
Next I compute the metrics for the future models. Like before I identify the data for the SSPs and Times combinations.
```{r}

SSPs <- c("ssp126", "ssp245","ssp370", "ssp585")
Times <- c(2050, 2085)


list_ssp <- 
  strsplit(names(chill_future_scenario_list), '\\.') %>%
  map(2) %>%
  unlist()

list_gcm <-
  strsplit(names(chill_future_scenario_list), '\\.') %>%
  map(3) %>%
  unlist()

list_time <-
  strsplit(names(chill_future_scenario_list), '\\.') %>%
  map(4) %>%
  unlist()



# now I loop through the different combinations and apply the workflow I used for the observed and past scenarios.

for(SSP in SSPs)
  for(Time in Times)
    {
    Temps <- future_temps[list_ssp == SSP & list_time == Time]
    names(Temps) <- list_gcm[list_ssp == SSP & list_time == Time]
    daily_models_future_scenarios <- tempResponse_list_daily(
      Temps,
      Start_JDay = Start_JDay,
      End_JDay = End_JDay,
      models = daily_models)
    daily_models_future_scenarios<-lapply(
      daily_models_future_scenarios,
      function(x) x[which(x$Perc_complete>90),])
    hourly_models_future_scenarios<-
      tempResponse_daily_list(
        Temps,
        latitude = 37.8667,
        Start_JDay = Start_JDay,
        End_JDay = End_JDay,
        models=hourly_models,
        misstolerance = 10)

    future_scenarios <- daily_models_future_scenarios
    
    future_scenarios <- lapply(
      names(future_scenarios),
      function(x)
        cbind(future_scenarios[[x]],
              hourly_models_future_scenarios[[x]][,names(hourly_models)]))
    names(future_scenarios)<-names(daily_models_future_scenarios)
    
    chill<-future_scenarios
    
    save_temperature_scenarios(
      chill,
      "data/future_climate",
      paste0("Dehesa_multichill_305_59_",Time,"_",SSP))
}
```
In the next step I load the created scenarios and use them to generate chill scenarios for plotting, using the make_climate_scenario()-function.

```{r}
chill_past_scenarios <- load_temperature_scenarios(
  "future_climate",
  "Dehesa_multichill_305_59_historic")

chill_observed <- read_tab("future_climate/Dehesa_multichill_305_59_observed.csv")

SSPs <- c("ssp126", "ssp245","ssp370", "ssp585")
Times <- c(2050, 2085)

# first I make the climate scenarios for the past scenarios and add the observed chill as historic data.
chills <- make_climate_scenario(chill_past_scenarios,
                                caption = "Historical",
                                historic_data = chill_observed,
                                time_series = TRUE)

# now I again loop through the SSP and time combinations for the future data and add the results to the list with the past and observed results.

for(SSP in SSPs)
  for(Time in Times)
    {
    chill <- load_temperature_scenarios(
      "future_climate",
      paste0("Dehesa_multichill_305_59_",Time,"_",SSP))
    if(SSP == "ssp126") SSPcaption <- "SSP1"
    if(SSP == "ssp245") SSPcaption <- "SSP2"
    if(SSP == "ssp370") SSPcaption <- "SSP3"
    if(SSP == "ssp585") SSPcaption <- "SSP5"    
    if(Time == "2050") Time_caption <- "2050"
    if(Time == "2085") Time_caption <- "2085"
    chills <- make_climate_scenario(chill,
                                    caption = c(SSPcaption,
                                                Time_caption),
                                    add_to = chills)
}
```
As stated before I want to crate heat maps with Safe Winter Chill metric for the models. For that I will use ggplot.

```{r}
# for that I first reorganize the data and safe it into a new data frame, to make it usable for ggplot. To compute Safe Winter Chill for the models and scenarios we use the quantile() function and use the results as input (quantile(ch$data[[nam]][,met],0.1))
for(i in 1:length(chills))
   {ch <- chills[[i]]
   if(ch$caption[1] == "Historical")
     {GCMs <- rep("none",length(names(ch$data)))
      SSPs <- rep("none",length(names(ch$data)))
      Years <- as.numeric(ch$labels)
      Scenario <- rep("Historical",
                      length(names(ch$data)))} else
                        {GCMs <- names(ch$data)
                        SSPs <- rep(ch$caption[1],
                                    length(names(ch$data)))
                        Years <- rep(as.numeric(ch$caption[2]),
                                     length(names(ch$data)))
                        Scenario <- rep("Future",
                                        length(names(ch$data)))}
   
   for(nam in names(ch$data))
     {for(met in metrics)
       {temp_res <-
         data.frame(Metric = met,
                    GCM = GCMs[which(nam == names(ch$data))],
                    SSP = SSPs[which(nam == names(ch$data))],
                    Year = Years[which(nam == names(ch$data))],
                    Result = quantile(ch$data[[nam]][,met],0.1), 
                    Scenario = Scenario[which(nam == names(ch$data))])
       if(i == 1 & nam == names(ch$data)[1] & met == metrics[1])
         results <- temp_res else
           results <- rbind(results,
                            temp_res)
         }
     }
   }

# In the end I normalize the results by expressing them as changes to the baseline year 1989, as the chill models use different units.
for(met in metrics)
  results[which(results$Metric == met),"SWC"] <-
    results[which(results$Metric == met),"Result"]/
      results[which(results$Metric == met & results$Year == 1989),
              "Result"]-1
```
The new created data.frame I can use now to plot the data, using ggplot. 
```{r}
rng = range(results$SWC)

# to plot the future results I plot all results where GCM is not equal to none

p_future <- ggplot(results[which(!results$GCM == "none"),],
                   aes(GCM,
                       y = factor(Metric,
                                  levels = metrics),
                       fill = SWC)) +
  geom_tile()

# to better see the single model results I rework the design of the plot
p_future <-
  p_future +
  facet_grid(SSP ~ Year) +
  theme_bw(base_size = 15) +
  theme(axis.text = element_text(size=6))+
  scale_fill_gradientn(colours = matlab.like(15),
                       labels = scales::percent,
                       limits = rng) +
  theme(axis.text.x = element_text(angle = 75, 
                                   hjust = 1,
                                   vjust = 1)) +
  labs(fill = "Change in\nSafe Winter Chill\nsince 1980") +
  scale_y_discrete(labels = model_labels) +
  ylab("Chill metric")

p_future
```


I also create plots for the past scenarios

```{r}
p_past<-
  ggplot(results[which(results$GCM == "none"),],
         aes(Year,
             y = factor(Metric, 
                        levels=metrics),
             fill = SWC)) +
  geom_tile()

p_past<-
  p_past +
  theme_bw(base_size = 15) +
  theme(axis.text = element_text(size = 6))

p_past<-
  p_past +
  scale_fill_gradientn(colours = matlab.like(20),
                       labels = scales::percent,
                       limits = rng)

p_past<-
  p_past +
  scale_x_continuous(position = "top") 

p_past<-
  p_past +
  labs(fill = "Change in\nSafe Winter Chill\nsince 1980") +
  scale_y_discrete(labels = model_labels) +
  ylab("Chill metric")

p_past
```

Now I combine them again using the patchwork package.

```{r}
chill_comp_plot<-
  (p_past +
     p_future +
     plot_layout(guides = "collect",
                 nrow = 2,
                 heights = c(1,3))) &
  theme(legend.position = "right",
        strip.background = element_blank(),
        strip.text = element_text(face = "bold"))

chill_comp_plot
```
The plot above shows the change in Safe Winter Chill at the Dehesa San Francisco since the year 1980, according to the different chill models. for the historic data, all models showed almost no change in SWC. For the future scenarios we can see mixed results. While most models showed an decrease between 0% and 150%, some model expect a decrease of up to 400%.

#3 Produce an animated line plot of your results (summarizing Safe Winter Chill across all the GCMs).

it seems also be interesting to know how the SWC developes for the different SSPs through out the years. For that we will produce an animated line plot.

```{r}
# first I duplicate the historic results and assign them to to the different SSP scenarios to easier plot the time series.
hist_results <- results[which(results$GCM == "none"),]
hist_results$SSP <- "SSP1"
hist_results_2 <- hist_results
hist_results_2$SSP <- "SSP2"
hist_results_3 <- hist_results
hist_results_3$SSP <- "SSP3"
hist_results_4 <- hist_results
hist_results_4$SSP <- "SSP5"
hist_results <- rbind(hist_results,
                      hist_results_2,
                      hist_results_3,
                      hist_results_4)

# now I summarize the results

future_results <- results[which(!results$GCM == "none"),]

GCM_aggregate <- aggregate(
  future_results$SWC,
  by=list(future_results$Metric,
          future_results$SSP,
          future_results$Year),
  FUN=mean)

colnames(GCM_aggregate) <- c("Metric",
                             "SSP",
                             "Year",
                             "SWC")

SSP_Time_series<-rbind(hist_results[,c("Metric",
                                       "SSP",
                                       "Year",
                                       "SWC")],
                       GCM_aggregate)
```
In the next step I create line plots showing the change for the SSP scenarios over time. 
```{r}
SSP_Time_series$Year <- as.numeric(SSP_Time_series$Year)

chill_change_plot<-
  ggplot(data = SSP_Time_series,
         aes(x = Year,
             y = SWC,
             col = factor(Metric,
                          levels = metrics))) +
  geom_line(lwd = 1.3) +
  facet_wrap(~SSP,
             nrow = 4) +
  theme_bw(base_size = 10) +
  labs(col = "Change in\nSafe Winter Chill\nsince 1980") +
  scale_color_discrete(labels = model_labels) +
  scale_y_continuous(labels = scales::percent) +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold")) +
  ylab("Safe Winter Chill")

chill_change_plot
```
Using the animate()- function to create an animated line plot.
```{r}
library(gganimate)
library(gifski)
library(png)
library(transformr)

ccp<-chill_change_plot +
  transition_reveal(Year)

animate(ccp, fps = 10)

anim_save("data/chill_comparison_animation.gif",
          animation = last_animation())
```

