---
output:
  pdf_document: default
  html_document: default
---
# Chapter 31 The PhenoFLex model - a second

```{r}
library(chillR)
library(tidyverse)
```


1. Make chill and heat response plots for the ‘Roter Boskoop’ PhenoFlex model for the location you did the earlier analyses for.

To evaluate the temperature response of the PhenoFlex components by producing temperature response plots I can again use the workflow developed in chapter 25. As the PhenoFlex model works in a different way as the other models I have to adept the code a bit. Instead of calculating chilling portions with the dynamic model I compute chilling efficiency with the PhenoFlex model. In addition I calculate the heat efficiency, weight by the fitted parameters. 

First I set up a function to weigh the GDH by the parameters Tb, Tu and Tc of the PhenoFlex Model.


```{r}


GDH_response <- function(T, par)
  {Tb <- par[11]
   Tu <- par[4]
   Tc <- par[10]
   GDH_weight <- rep(0, length(T))
   GDH_weight[which(T >= Tb & T <= Tu)] <-
     1/2 * (1 + cos(pi + pi * (T[which(T >= Tb & T <= Tu)] - Tb)/(Tu - Tb)))
   GDH_weight[which(T > Tu & T <= Tc)] <-
     (1 + cos(pi/2 + pi/2 * (T[which(T >  Tu & T <= Tc)] -Tu)/(Tc - Tu)))
  return(GDH_weight)
}
```

Now I make the described changes to the function of chapter 25 and run it.

```{r,eval=FALSE}

latitude <- 50.6

month_range <- c(10, 11, 12, 1, 2, 3)

Tmins = c(-20:20)
Tmaxs = c(-15:30)

mins <- NA
maxs <- NA
chill_eff <- NA
heat_eff <- NA
month <- NA

simulation_par <- read_tab("data/PhenoFlex_parameters_Boskop.csv")[,2]

for(mon in month_range)
    {days_month <- as.numeric(difftime(ISOdate(2002, mon+1, 1),
                                       ISOdate(2002, mon, 1)))
     if(mon == 12) days_month <- 31
     weather <- 
       make_all_day_table(data.frame(Year = c(2002, 2002),                                   
                                     Month = c(mon, mon),
                                     Day = c(1, days_month),
                                     Tmin = c(0, 0),
                                     Tmax = c(0, 0)))
     
     for(tmin in Tmins)
      for(tmax in Tmaxs)
        if(tmax >= tmin)
          {
           hourtemps <- weather %>%
             mutate(Tmin = tmin,
                    Tmax = tmax) %>%
             stack_hourly_temps(latitude = latitude) %>%
             pluck("hourtemps", "Temp")
           
           chill_eff <- 
             c(chill_eff,
               PhenoFlex(temp = hourtemps,
                         times = c(1: length(hourtemps)),
                         A0 = simulation_par[7],
                         A1 = simulation_par[8],
                         E0 = simulation_par[5],
                         E1 = simulation_par[6],
                         Tf = simulation_par[9],
                         slope = simulation_par[12],
                         deg_celsius = TRUE,
                         basic_output = FALSE)$y[length(hourtemps)] /
                                            (length(hourtemps) / 24))
           
          heat_eff <- 
            c(heat_eff,
              cumsum(GDH_response(hourtemps,
                                  simulation_par))[length(hourtemps)] /
                                                 (length(hourtemps) / 24))
          mins <- c(mins, tmin)
          maxs <- c(maxs, tmax)
          month <- c(month, mon)
        }
}

results <- data.frame(Month = month,
                      Tmin = mins,
                      Tmax = maxs,
                      Chill_eff = chill_eff,
                      Heat_eff = heat_eff) %>%
  filter(!is.na(Month))

#saving the results

write.csv(results,
          "data/model_sensitivity_PhenoFlex.csv")
```




To plot the chill & heat efficiencies I use the Chill_sensitivity_temps function again, developed in chapter 25.

```{r,eval=FALSE}
Model_sensitivities_PhenoFlex <-
  read.csv("data/model_sensitivity_PhenoFlex.csv")

CKA_weather <- read_tab("data/TMaxTMin1958-2019_patched.csv")


Chill_sensitivity_temps(Model_sensitivities_PhenoFlex,
                        CKA_weather,
                        temp_model = "Chill_eff",
                        month_range = c(10, 11, 12, 1, 2, 3),
                        Tmins = c(-20:20),
                        Tmaxs = c(-15:30),
                        legend_label = "Chill per day \n(arbitrary)") +
  ggtitle("PhenoFlex chill efficiency ('Roter Boskoop')")
```

```{r,eval=FALSE}
Chill_sensitivity_temps(Model_sensitivities_PhenoFlex,
                        CKA_weather,
                        temp_model = "Heat_eff",
                        month_range = c(10, 11, 12, 1, 2, 3),
                        Tmins = c(-20:20),
                        Tmaxs = c(-15:30),
                        legend_label = "Heat per day \n(arbitrary)") +
  ggtitle("PhenoFlex heat efficiency ('Roter Boskoop')")
```

